---
title: "Seminar Neuro - Modelfit"
author: "28.03.2017"
date: "30 April 2017"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Modelfit
## Daten 

```{r experiment}
values <- reactiveValues()
```


```{r df, echo=FALSE}

df1 = data.frame(Gewinn_Jetzt=rep(c(200,150,100,80,30,10),5),
                 Gewinn_Spaeter=rep(200,30),
                 delay_Wochen=rep(c(2,4,8,16,32),each=6))


values$delay <- unique(df1$delay_Wochen)

# DT::renderDataTable(DT::datatable(
#   df1, options = list(searching = FALSE, 
#                          paging = FALSE)
# ))
```

copy&paster  80,100,120,140,180

```{r inputName}
inputPanel(
textInput("namePerson1", "Person 1", value = "Person 1"),
textInput("namePerson2", "Person 2", value = "Person 2")
)
```

```{r inputExp}

observe({
  n1          <- input$namePerson1
  updateTextInput(session,"inputExp1", n1, value = "")
  
  
  n2         <- input$namePerson2
  updateTextInput(session,"inputExp2", n2, value = "")
 
})


inputPanel(
  textInput("inputExp1", "Person1", value = ""),
  textInput("inputExp2", "Person2", value = "")
)
```

##  Plot:Indifferenzpunkte ~ Delay



```{r plot raw values}

#   
#   indPointP1 <- as.numeric(unlist(strsplit(input$inputExp1, ",")))
#   indPointP2 <- as.numeric(unlist(strsplit(input$inputExp2, ","))) 
#   
#   
#   

renderPlot({
  
  values$indPointP1 <- as.numeric(unlist(strsplit(input$inputExp1, ",")))
  values$indPointP2 <- as.numeric(unlist(strsplit(input$inputExp2, ","))) 
  
  
 
  par(mfrow=c(1,2))
  plot(values$indPointP1~values$delay,
       ylab="Wert Indifferenzpunkte",xlab="Delay in Wochen",main=input$namePerson1,cex.lab=1.3,cex = 2.5,pch=16)
  
  plot(values$indPointP2~values$delay,
       ylab="Wert Indifferenzpunkte",xlab="Delay in Wochen",main=input$namePerson2,cex.lab=1.3,cex = 2.5,pch=16)
  
})


```

## Modelfit
### Bestimmung der Funktion 

Hier verwenden wir die die hyperbolische Funktion mit der Funktionsgleichung:
$I=\frac{A}{(1+k*D)}$ wobei $I$ =  der  Wert des Indifferenzpunktes f?r den Betrag $A$ nach einer Zeit von $D$ ist. 

Beispiel:

Code:

```{r, eval=FALSE, echo=TRUE}

 hyperbolic<- function(D,A,k){
      A/(1+k*D)
    }
    
    plot(y ~ x,cex = 2.5,pch=16)
    curve(hyperbolic(x,1000,0.2),add=TRUE,col="red",lwd=5)
```

Wie sieht funktion aus, wenn sich k zu 0.3 oder 0.01 ?ndert?

```{r}
numericInput("k", "k:", 0.2)

hyperbolic<- function(D,A,k){
      A/(1+k*D)
    }

renderPlot({
    x  <- c(1,2,4,25,52,156,520)
    y  <- c(750,700,450,250,100,60,60)
    
    
    plot(y ~ x,cex = 2.5,pch=16)
    curve(hyperbolic(x,1000,input$k),add=TRUE,col="red",lwd=5)
 })
```


## Bestimmt der likelihood f?r einen bestimmten Parameter Wert

```{r}
h.loss <- function(par,x,y) {
  
  k.par     <-  par[1] 
  err.sigma <-  par[2]  
  
# If the error standard deviation is invalid (i.e.; negative), then we need to return a very high deviance
  
# This will tell the optimization procedure to stay away from invalid (either statistically or psychologically) parameter values.
  
  if(err.sigma < 0) {deviance <- 10000000}
  
  # If the error standard deviation is valid (i.e.; > 0), then calculate the deviance...
  
  if(err.sigma > 0) {
    
    # Calculate the likelihood of each data point.
    # Here is where you specify the model and how you calculate likelihoods.
    
    likelihoods <- dnorm(y, mean = hyperbolic(x,1000,k.par), sd = err.sigma)
    
    # Now let's convert the vector of likelihoods to a summary deviance score (-2 times sub log-lik)
    
    # Calculate log-likelihoods of each data point
    log.likelihoods <- log(likelihoods)
    
    # Calculate the deviance (-2 times sum of log-likelihoods)
    deviance <- -2 * sum(log.likelihoods)
    
  }
  
  return(deviance)
  
}
```

Beispiel - Daten:

    Delay  <- c(1,2,4,25,52,156,520)
    Indifferenzwert  <- c(750,700,450,250,100,60,60)
    k=0.2

F?r den ersten Datenpunkt w?re die likelihood dann:

```{r, eval=FALSE, echo=TRUE}
likelihood <- dnorm(750, mean = hyperbolic(1,1000,0.2), sd = 10)
```

```{r}

    x  <- c(1,2,4,25,52,156,520)
    y  <- c(750,700,450,250,100,60,60)
    
    
(likelihoods <- dnorm(750, mean = hyperbolic(1,1000,0.2), sd = 58))
```

und f?r alle:

```{r}


    x  <- c(1,2,4,25,52,156,520)
    y  <- c(750,700,450,250,100,60,60)
    
    
likelihoods <- dnorm(y, mean = hyperbolic(x,1000,0.2), sd = 58)

plot(likelihoods~x,ylab="Likelihoods",xlab="Delay",cex.lab=1.2,ylim=c(-1,1),col="red",lwd=2,cex = 2.5,pch=16)
```


Deswegen verwendet man die log(LL):

```{r}
plot(log(likelihoods)~x,ylab="Likelihoods",xlab="Delay",cex.lab=1.2,col="red",lwd=2)
```


Nun summiert man die log-LL Werte zusammen und erh?lt so die log(LL) oder die Devianz = -2*log(LL) f?r den k=0.2

```{r}
sum(log(likelihoods))
sum(log(likelihoods))*-2
```

## Optimierung

Jetzt setzt man iterativ verschieden Werte f?r k ein und bestimmt den Wert mit der h?chsten log(LL):

```{r}
sliderInput("k2", "k:", min = 0.01, max = 1, value = 0.01, step= 0.05,
            animate = animationOptions(interval=1000))

renderPlot({

 
    x  <- c(1,2,4,25,52,156,520)
    y  <- c(750,700,450,250,100,60,60)
    
    
likelihoods <- dnorm(y, mean = hyperbolic(x,1000,input$k2), sd = 58)



plot(1, type="n" ,ylab="log(LL)",xlab="k",cex.lab=1.2,xlim=c(0, 1), ylim=c(-500,0))
points(y=h.loss(c(input$k2,58),x,y)*-2,x=input$k2,col="red",lwd=5)

})

dev.temp <- c()


for(i in 1:length(seq(0.01,1,0.03))){
  dev.temp[i] <- h.loss(c(seq(0.01,1,0.03)[i],58),x,y)*-2
}


plot(1, type="n" ,ylab="log(LL)",xlab="k",cex.lab=1.2,xlim=c(0, 1), ylim=c(-500,0))
for(i in seq(0.01,1,0.03)){
  dev.temp1 <- h.loss(c(i,58),x,y)*-2
  points(x=i,y=dev.temp1)
  
}


abline(v = seq(0.01,1,0.03)[dev.temp==max(dev.temp)])





```


## Sch?tzung von k f?r unsere Daten


```{r}

renderText({
  
  #x <- c(2,4,8,16,32)
  #y <- c(150,100,60,60,30)
  
  h.loss1 <- function(par) {
  
  k.par     <-  par[1] 
  err.sigma <-  par[2]  
  
# If the error standard deviation is invalid (i.e.; negative), then we need to return a very high deviance
  
# This will tell the optimization procedure to stay away from invalid (either statistically or psychologically) parameter values.
  
  if(err.sigma < 0) {deviance <- 10000000}
  
  # If the error standard deviation is valid (i.e.; > 0), then calculate the deviance...
  
  if(err.sigma > 0) {
    
    # Calculate the likelihood of each data point.
    # Here is where you specify the model and how you calculate likelihoods.
    
    likelihoods <- dnorm(values$indPointP1, mean = hyperbolic(values$delay,200,k.par), sd = err.sigma)
    
    # Now let's convert the vector of likelihoods to a summary deviance score (-2 times sub log-lik)
    
    # Calculate log-likelihoods of each data point
    log.likelihoods <- log(likelihoods)
    
    # Calculate the deviance (-2 times sum of log-likelihoods)
    deviance <- -2 * sum(log.likelihoods)
    
  }
  
  return(deviance)
  
  }
  
  

  
  
  parameter.fitsP1 <- optim(par = c(0.5, 100),
                          fn = h.loss1,
                          hessian = T
)
  
values$kP1 <- parameter.fitsP1$par[1]

  
})



```

```{r}

renderText({
h.loss2 <- function(par) {
  
  k.par     <-  par[1] 
  err.sigma <-  par[2]  
  
# If the error standard deviation is invalid (i.e.; negative), then we need to return a very high deviance
  
# This will tell the optimization procedure to stay away from invalid (either statistically or psychologically) parameter values.
  
  if(err.sigma < 0) {deviance <- 10000000}
  
  # If the error standard deviation is valid (i.e.; > 0), then calculate the deviance...
  
  if(err.sigma > 0) {
    
    # Calculate the likelihood of each data point.
    # Here is where you specify the model and how you calculate likelihoods.
    
    likelihoods <- dnorm(values$indPointP2, mean = hyperbolic(values$delay,200,k.par), sd = err.sigma)
    
    # Now let's convert the vector of likelihoods to a summary deviance score (-2 times sub log-lik)
    
    # Calculate log-likelihoods of each data point
    log.likelihoods <- log(likelihoods)
    
    # Calculate the deviance (-2 times sum of log-likelihoods)
    deviance <- -2 * sum(log.likelihoods)
    
  }
  
  return(deviance)
  
}


    parameter.fitsP2 <- optim(par = c(0.5, 100),
                          fn = h.loss2,
                          hessian = T
)

values$kP2 <- parameter.fitsP2$par[1]
    
    
   }) 
```



```{r}
renderPlot({ 
  
  par(mfrow=c(1,2))
  x <- values$delay

  plot(values$indPointP1 ~ values$delay,ylim=c(0,200),xlim=c(0,32),ylab="Wert Indifferenzpunkt",xlab="Delay",cex.lab=1.3,cex = 2.5,pch=16)
  curve(hyperbolic(x,200,values$kP1),add=TRUE,col="red",lwd=5)
  text(25, 100, values$kP1, cex = 1.8)

  plot(values$indPointP2 ~ values$delay,ylim=c(0,200),xlim=c(0,32),ylab="Wert Indifferenzpunkt",xlab="Delay",cex.lab=1.3,cex = 2.5,pch=16)
  curve(hyperbolic(x,200,values$kP2),add=TRUE,col="red",lwd=5)
  text(25, 100, values$kP2, cex = 1.8)
  })
```

